---
title: "dsh_analysis_R"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)

dsh_2011 <- read_csv('dsh_2011.csv')
dsh_2013 <- read_csv('dsh_2013.csv')
dsh_2015 <- read_csv('dsh_2015.csv')
states <- read.csv('expansion_states.csv')

dsh_2011[is.na(dsh_2011)] <- 0
dsh_2013[is.na(dsh_2013)] <- 0
dsh_2015[is.na(dsh_2015)] <- 0
```

1) Find the total DSH payments each state receives in total in 2011 (aggregated from all hospitals for the state).

2) Find the numbers for year 2015.

3) Find how much DSH payments has increased in each state ((2) - (1)).

4) Label each state as "expansion states" or "non-expansion states", according to whether they decided to expand Medicaid following the Affordable Care Act. 

5) Use a logistic regression model to describe the relation between DSH increase and expansion/non-expansion. Is there a strong correlation? If so, it may support the argument that Medicaid expansion helped reduce the DSH payments 
```{r}
dsh_2011 <-
  dsh_2011 %>%
  select(state, dsh1, dsh2, dsh3)

dsh_2011 <-
  dsh_2011 %>%
  group_by(state) %>%
  summarize_all(sum)

dsh_2013 <-
  dsh_2013 %>%
  select(state, dsh1, dsh2, dsh3)

dsh_2013 <-
  dsh_2013 %>%
  group_by(state) %>%
  summarize_all(sum)

dsh_2015 <-
  dsh_2015 %>%
  select(state, dsh1, dsh2, dsh3)

dsh_2015$dsh3 <- as.numeric(dsh_2015$dsh3)

dsh_2015 <-
  dsh_2015 %>%
  group_by(state) %>%
  summarize_all(sum)
```

```{r}
dsh_2011$dsh11 <-
  dsh_2011$dsh1 + dsh_2011$dsh2 + dsh_2011$dsh3

dsh_2011 <-
  dsh_2011 %>%
  select(state, dsh11)

dsh_2013$dsh13 <-
  dsh_2013$dsh1 + dsh_2013$dsh2 + dsh_2013$dsh3

dsh_2013 <-
  dsh_2013 %>%
  select(state, dsh13)

dsh_2015$dsh15 <-
  dsh_2015$dsh1 + dsh_2015$dsh2 + dsh_2015$dsh3

dsh_2015 <-
  dsh_2015 %>%
  select(state, dsh15)

```

```{r}
dsh <- merge(dsh_2011, dsh_2015, by = 'state')
dsh <- merge(dsh, dsh_2013, by = 'state')
```

```{r}
dsh$x15to11 <-
  (dsh$dsh15 - dsh$dsh11)/(dsh$dsh11)

dsh$x13to11 <-
  (dsh$dsh13 - dsh$dsh11)/(dsh$dsh11)
```

```{r}
dsh_states <- merge(dsh, states, by = 'state')
```

```{r}
model <- glm(expansion_status ~ x15to11, data = dsh_states, family = 'binomial')

xweight <- data.frame(x15to11 = seq(min(dsh_states$x15to11), max = dsh_states$x15to11), len = 100)

yweight <- predict(model, list(expansion_status = xweight), type = "response")

plot(dsh_states$expansion_status ~ dsh_states$x15to11, col = "red4")

lines(xweight, yweight)
```





